Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}

### 2021-12-12 article-knitr.tex
update paper and plots

\begin{figure}[h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=0.98\textwidth]{/mnt/data/GoogleDrive/Uni/PhD/scoringutils-paper/plots/calibration-diagnostic-examples} 
\end{knitrout}

@@

\subsection{Preparing and visualising the data}

We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
% 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# install and load data pacakge from external repository}
\hlcom{# remotes::install_github('sbfnk/covid19.forecasts.uk')}

\hlcom{# load packages}
\hlkwd{library}\hlstd{(covid19.forecasts.uk)}
\hlkwd{library}\hlstd{(dplyr)}
\hlkwd{library}\hlstd{(scoringutils)}

\hlcom{# load truth data}
\hlkwd{data}\hlstd{(covid_uk_data)}
\hlcom{# head(covid_uk_data)}

\hlcom{# load forecasts}
\hlkwd{data}\hlstd{(uk_forecasts)}
\hlcom{# head(uk_forecasts)}
\end{alltt}
\end{kframe}
\end{knitrout}
% 
Let us take a first look a the data:
% 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{glimpse}\hlstd{(covid_uk_data)}
\end{alltt}
\begin{verbatim}
## Rows: 4,174
## Columns: 6
## $ geography  <fct> London, London, London, London, London, London, L~
## $ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita~
## $ value_desc <fct> Hospital admissions, Hospital admissions, Hospita~
## $ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0~
## $ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, ~
## $ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, ~
\end{verbatim}
\begin{alltt}
\hlkwd{glimpse}\hlstd{(uk_forecasts)}
\end{alltt}
\begin{verbatim}
## Rows: 1,254,513
## Columns: 8
## $ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E~
## $ geography     <chr> "East of England", "East of England", "East of~
## $ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp~
## $ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3~
## $ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0~
## $ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05~
## $ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0~
## $ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp~
\end{verbatim}
\end{kframe}
\end{knitrout}
% 
Some minor changes need to be made to the data before we can evaluate the forecasts. The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}. While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public. The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way. 
% 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{uk_forecasts} \hlkwb{<-} \hlkwd{rename}\hlstd{(uk_forecasts,} \hlkwc{prediction} \hlstd{= value)}
\hlstd{covid_uk_data} \hlkwb{<-} \hlkwd{rename}\hlstd{(covid_uk_data,} \hlkwc{true_value} \hlstd{= value)}

\hlstd{combined} \hlkwb{<-} \hlkwd{merge_pred_and_obs}\hlstd{(uk_forecasts, covid_uk_data)}
\end{alltt}
\end{kframe}

@@

% 
It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance. The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts. 
% 
\begin{figure}[h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{show_avail_forecasts}\hlstd{(}\hlkwc{data} \hlstd{= combined,} \hlkwc{x} \hlstd{=} \hlstr{"creation_date"}\hlstd{,} \hlkwc{show_numbers} \hlstd{=} \hlnum{FALSE}\hlstd{,}
    \hlkwc{legend_position} \hlstd{=} \hlstr{"bottom"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc)}
\end{alltt}
\end{kframe}
\includegraphics[width=0.98\textwidth]{plots/plot-show-availability-1} 
\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function. Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases. In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values. This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately. Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: % 
\begin{figure}[h!]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{locations} \hlkwb{<-} \hlstr{"c(\textbackslash{}"Scotland\textbackslash{}", \textbackslash{}"Wales\textbackslash{}", \textbackslash{}"Northern Ireland\textbackslash{}", \textbackslash{}"England\textbackslash{}")"}
\hlkwd{plot_predictions}\hlstd{(}\hlkwc{truth} \hlstd{= covid_uk_data,} \hlkwc{forecasts} \hlstd{= uk_forecasts,} \hlkwc{filter_both} \hlstd{=} \hlkwd{list}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlstr{"geography %in%"}\hlstd{,}
    \hlstd{locations,} \hlkwc{collapse} \hlstd{=} \hlstr{""}\hlstd{),} \hlstr{"value_type == \textbackslash{}"death_inc_line\textbackslash{}""}\hlstd{),} \hlkwc{filter_truth} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"value_date <= \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{,}
    \hlstr{"value_date > \textbackslash{}"2020-06-01\textbackslash{}""}\hlstd{),} \hlkwc{filter_forecasts} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"model == \textbackslash{}"SIRCOVID\textbackslash{}""}\hlstd{,}
    \hlstr{"creation_date == \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{),} \hlkwc{x} \hlstd{=} \hlstr{"value_date"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{geography)} \hlopt{+}
    \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
\end{alltt}
\end{kframe}
\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

@@

The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}. The function automatically detects the forecast type and format, applies the appropriate scoring metrics and aggregates results as desired by the user. Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation. The function requires a \code{data.frame} or similar with some pre-specified columns (e.g. a column called "prediction", one called "true\_value" and more depending on the forecast format). Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons. To illustrate the requirements, \pkg{scoringutils} provides example data for all possible forecast formats. It also offers functions to to transform between various formats, e.g. from a sample based format to a quantile format.

When conducting an evaluation, we need to decide on the level of grouping over which we want to aggregate scores. For the UK short-term forecasts it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
% ion G
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"value_desc"}\hlstd{))}

\hlkwd{glimpse}\hlstd{(scores)}
\end{alltt}
\begin{verbatim}
## Rows: 29
## Columns: 9
## $ model              <fct> EpiSoon, StructuredODE, Microsimulation, ~
## $ value_desc         <fct> Hospital admissions, Hospital admissions,~
## $ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876, ~
## $ dispersion         <dbl> 22.886153, 6.133974, 33.985553, 10.365082~
## $ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.5~
## $ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.55~
## $ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174, ~
## $ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0.~
## $ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563, ~
\end{verbatim}
\end{kframe}
\end{knitrout}
% 
\fct{eval\_forecasts} aggregates scores by taking the mean over according to the grouping specified in \code{summarise_by}. In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target. For a more detailed analysis we could for example specifiy \code{summarise_by = c("mode", "value_type", "geography")} to additionally stratify by location. If we wanted to have one score per quantile or one per prediction interval range, we could use something like \code{summarise_by = c("model", "quantile", "range")}. This would allow us to examine interval or quantile coverage or makes it possible to analyse the accuracy of the tails of the forecasts. In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.

The user must, however, exercise some caution when aggregating scores. As explained above, many of the metrics are absolute and scale with the magnitude of the quantity to forecast. This makes it sometimes ill-advised to average over them. In the given example, looking at one score per model (i.e. specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence. Similarly, aggregating over different forecast horizons is often not a good idea as the mean will be dominated by further ahead forecast horizons.

The second important argument of the \fct{eval\_forecasts} function is the \code{by} argument that denotes the unit of a single forecast. In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")} as there is one distinct forecast for each combination of these categories. Quantiles do not need to be included, as several quantiles make up one forecast (and similarly for samples). By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast. In the above example, we did not have to explicitly specify the \code{by} argument, as all available columns were relevant to the unit of a single forecast. If there are additional columns that do not correspond to the unit of a single forecast, like for example the observed temperature in a given location, \code{by} must be specified and these columns must be excluded.

@@

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# unsummarised scores}
\hlstd{unsummarised_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined)}
\hlstd{pairwise} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(unsummarised_scores,} \hlkwc{summarise_by} \hlstd{=} \hlstr{"value_desc"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

>>>>

\begin{figure}[h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{/mnt/data/GoogleDrive/Uni/PhD/scoringutils-paper/plots/calibration-diagnostic-examples} 
\end{knitrout}

@@

\subsection{Checking the data}

% 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlcom{# load packages}
\hlstd{> }\hlkwd{library}\hlstd{(scoringutils)}
\hlstd{> }\hlkwd{library}\hlstd{(dplyr)}
\hlstd{> }\hlkwd{library}\hlstd{(data.table)}
\hlstd{> }\hlkwd{library}\hlstd{(kableExtra)}
\hlstd{> }
\hlstd{> }\hlcom{# load forecasts}
\hlstd{> }\hlkwd{data}\hlstd{(example_quantile)}
\end{alltt}
\end{kframe}
\end{knitrout}
% 
All higher level functions that allow for convenient evaluation of forecasts expect a \code{data.frame} or similar. Lower-level functions (discussed in a later section) accept inputs as vectors and matrices.

We omit \code{NA} values as the data set also contains entries for which we have an observed value, but no forecast. 
% 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{example_quantile |>}
\hlstd{+ }    \hlkwd{na.omit}\hlstd{() |>}
\hlstd{+ }    \hlkwd{glimpse}\hlstd{()}
\end{alltt}
\begin{verbatim}
## Rows: 20,401
## Columns: 10
## $ location        <chr> "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", ~
## $ target_end_date <date> 2021-05-08, 2021-05-08, 2021-05-08, 2021-05-08, 2021-~
## $ target_type     <chr> "Cases", "Cases", "Cases", "Cases", "Cases", "Cases", ~
## $ true_value      <dbl> 106987, 106987, 106987, 106987, 106987, 106987, 106987~
## $ location_name   <chr> "Germany", "Germany", "Germany", "Germany", "Germany",~
## $ forecast_date   <chr> "2021-05-03", "2021-05-03", "2021-05-03", "2021-05-03"~
## $ quantile        <dbl> 0.010, 0.025, 0.050, 0.100, 0.150, 0.200, 0.250, 0.300~
## $ prediction      <int> 82466, 86669, 90285, 95341, 99171, 102990, 105962, 108~
## $ model           <chr> "EuroCOVIDhub-ensemble", "EuroCOVIDhub-ensemble", "Eur~
## $ horizon         <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
\end{verbatim}
\end{kframe}
\end{knitrout}
%

For forecasts in a quantile-based format it is necessary to have columns called \code{true_value}, \code{prediction} and \code{quantile}. It is recommended (although not necessary except in order to do pairwise comparisons) to have a column called "model" with an identifier for the forecaster.
Table \ref{tab:column-requirements} shows the columns expected for different input formats.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{requirements} \hlkwb{<-} \hlkwd{data.table}\hlstd{(}\hlkwc{Format} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"quantile-based"}\hlstd{,} \hlstr{"sample-based"}\hlstd{,}
\hlstd{+ }    \hlstr{"binary"}\hlstd{,} \hlstr{"pairwise-comparisons"}\hlstd{),} \hlkwc{`Required columns`} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"'true_value', 'prediction', 'quantile'"}\hlstd{,}
\hlstd{+ }    \hlstr{"'true_value', 'prediction', 'sample'"}\hlstd{,} \hlstr{"'true_value', 'prediction'"}\hlstd{,}
\hlstd{+ }    \hlstr{"additionally a column 'model'"}\hlstd{))}
\hlstd{> }
\hlstd{> }\hlstd{requirements |>}
\hlstd{+ }    \hlkwd{kable}\hlstd{(}\hlkwc{format} \hlstd{=} \hlstr{"latex"}\hlstd{)}
\end{alltt}
\end{kframe}
\begin{tabular}{l|l}
\hline
Format & Required columns\\
\hline
quantile-based & 'true\_value', 'prediction', 'quantile'\\
\hline
sample-based & 'true\_value', 'prediction', 'sample'\\
\hline
binary & 'true\_value', 'prediction'\\
\hline
pairwise-comparisons & additionally a column 'model'\\
\hline
\end{tabular}

@@

We can check whether the data conforms to the requirements by running \fct{check\_forecasts}.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{check_forecasts}\hlstd{(example_quantile)}
\end{alltt}
\begin{verbatim}
## $target_type
## [1] "integer"
## 
## $prediction_type
## [1] "quantile"
## 
## $forecast_unit
## [1] "location"        "target_end_date" "target_type"     "location_name"  
## [5] "forecast_date"   "model"           "horizon"        
## 
## $unique_values
##                    model location target_end_date target_type true_value
## 1: EuroCOVIDhub-ensemble        4              12           2         96
## 2: EuroCOVIDhub-baseline        4              12           2         96
## 3:  epiforecasts-EpiNow2        4              12           2         95
## 4:       UMass-MechBayes        4              12           1         48
##    location_name forecast_date quantile prediction horizon
## 1:             4            11       23       3969       3
## 2:             4            11       23       3733       3
## 3:             4            11       23       3903       3
## 4:             4            11       23       1058       3
## 
## $warnings
## [1] "Some values for `prediction` are NA in the data provided"
## 
## 
## Based on your input, scoringutils thinks:
## Forecasts are for a `integer` target using a `quantile` prediction format.
## The unit of a single forecast is defined by `location`, `target_end_date`, `target_type`, `location_name`, `forecast_date`, `model`, `horizon`. If this is not as intended, please DELETE UNNECESSARY columns or add new ones.
## $unique_values shows how many unique values there are per column per model (across the entire data).
## 
## You should be aware of the following warnings:
## Some values for `prediction` are NA in the data provided
\end{verbatim}
\end{kframe}
\end{knitrout}

This returns a list with different entries giving information about what \pkg{scoringutils} infers from the data. \code{target_type} \code{prediction_type} refer to the forecast format. \code{forecast_unit} contains a vector of the columns which \pkg{scoringutils} thinks denote the unit of a single forecast. This means that in this instance a single forecast (with a set of 23 quantiles) can uniquely be identified by the values in the columns "location", "target\_end\_date", "target\_type", "location\_name", "forecast\_date", "model", "horizon". In this example, having "location" as well as "location\_name" included does not make a difference, as they contain duplicated information. In general, however, it is strongly advised to remove all unnecessary columns that do not help identify a single forecast. \code{unique_values} gives an overview of the number of unique values per column across the entire data set, providing a first hint as to whether the forecast set is complete. \code{warnings} shows potential warnings about the data. In this example, \pkg{scoringutils} warns that there are observed values present for which there is no corresponding forecast. These warnings can often be ignored, but may provide important information. If there are errors that cannot be ignored, a list entry \code{errors} will appear.

It is helpful to start the evaluation process by visualising data availability, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance. The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts. 
% 
\begin{figure}[h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{show_avail_forecasts}\hlstd{(}\hlkwc{data} \hlstd{= example_quantile,} \hlkwc{x} \hlstd{=} \hlstr{"target_end_date"}\hlstd{,}
\hlstd{+ }    \hlkwc{show_numbers} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{legend_position} \hlstd{=} \hlstr{"bottom"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{plots/plot-show-availability-1} 
\end{knitrout}
\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function. Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases. In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values. This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately. Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To display, for example, short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021 as well as 5 weeks of prior data, we can call: % 
\begin{figure}[h!]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{plot_predictions}\hlstd{(}\hlkwc{data} \hlstd{= example_quantile,} \hlkwc{x} \hlstd{=} \hlstr{"target_end_date"}\hlstd{,}
\hlstd{+ }    \hlkwc{filter_truth} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"target_end_date <= \textbackslash{}"2021-07-15\textbackslash{}""}\hlstd{,}
\hlstd{+ }        \hlstr{"target_end_date > \textbackslash{}"2021-05-22\textbackslash{}""}\hlstd{),} \hlkwc{filter_forecasts} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"model == 'EuroCOVIDhub-ensemble'"}\hlstd{,}
\hlstd{+ }        \hlstr{"forecast_date == \textbackslash{}"2021-06-28\textbackslash{}""}\hlstd{),} \hlkwc{facet_formula} \hlstd{= target_type} \hlopt{~}
\hlstd{+ }        \hlstd{location,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{plots/plot-show-forecasts-1} 
\end{knitrout}
\caption{\label{fig:forecast-visualisation} Short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

@@

The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}. The function automatically detects the forecast type and format (as shown in the output of \fct{check\_forecasts}), applies the appropriate scoring metrics and can in addition also aggregate results as well as perform pairwise comparisons between models. Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile)}
\hlstd{> }\hlkwd{glimpse}\hlstd{(scores)}
\end{alltt}
\begin{verbatim}
## Rows: 887
## Columns: 14
## $ location           <chr> "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE~
## $ target_end_date    <date> 2021-05-08, 2021-05-08, 2021-05-08, 2021-05-08, 20~
## $ target_type        <chr> "Cases", "Cases", "Cases", "Deaths", "Deaths", "Dea~
## $ location_name      <chr> "Germany", "Germany", "Germany", "Germany", "German~
## $ forecast_date      <chr> "2021-05-03", "2021-05-03", "2021-05-03", "2021-05-~
## $ model              <chr> "EuroCOVIDhub-baseline", "EuroCOVIDhub-ensemble", "~
## $ horizon            <dbl> 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 1, 1, 1, 2, 2, 2, 2, ~
## $ interval_score     <dbl> 16925.04696, 7990.85478, 25395.96087, 46.79304, 53.~
## $ dispersion         <dbl> 1649.22087, 5440.98522, 8173.70000, 44.66261, 53.27~
## $ underprediction    <dbl> 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.60869~
## $ overprediction     <dbl> 15275.826087, 2549.869565, 17222.260870, 2.130435, ~
## $ coverage_deviation <dbl> -0.38521739, 0.04956522, -0.29826087, 0.22347826, 0~
## $ bias               <dbl> 0.95, 0.50, 0.90, 0.30, -0.10, -0.50, 0.10, 1.00, 0~
## $ aem                <dbl> 25620, 12271, 44192, 15, 14, 208, 24, 67622, 45731,~
\end{verbatim}
\end{kframe}
\end{knitrout}

The above produces one score for every forecast. However, we usually like to summarise scores to learn about average performance.

We can either run a separate function to summarise the scores, 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{summarised_scores} \hlkwb{<-} \hlkwd{summarise_scores}\hlstd{(}\hlkwc{scores} \hlstd{= scores,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
\hlstd{+ }    \hlstr{"target_type"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

or we score and summarise in one function call:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{summarised_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
\hlstd{+ }    \hlstr{"target_type"}\hlstd{))}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{summarised_scores}
\end{alltt}
\begin{verbatim}
##                    model target_type interval_score dispersion underprediction
## 1: EuroCOVIDhub-baseline       Cases    28483.57465 4102.50094    10284.972826
## 2: EuroCOVIDhub-ensemble       Cases    17943.82383 3663.52458     4237.177310
## 3:  epiforecasts-EpiNow2       Cases    20831.55662 5664.37795     3260.355639
## 4: EuroCOVIDhub-baseline      Deaths      159.40387   91.40625        2.098505
## 5: EuroCOVIDhub-ensemble      Deaths       41.42249   30.18099        4.103261
## 6:       UMass-MechBayes      Deaths       52.65195   26.87239       16.800951
## 7:  epiforecasts-EpiNow2      Deaths       66.64282   31.85692       15.893314
##    overprediction coverage_deviation        bias         aem
## 1:   14096.100883        -0.11211957  0.09796875 38473.60156
## 2:   10043.121943        -0.09785326 -0.05640625 24101.07031
## 3:   11906.823030        -0.06660326 -0.07890625 27923.81250
## 4:      65.899117         0.11614130  0.33906250   233.25781
## 5:       7.138247         0.19528533  0.07265625    53.13281
## 6:       8.978601        -0.02312500 -0.02234375    78.47656
## 7:      18.892583        -0.04287176 -0.00512605   104.74790
\end{verbatim}
\end{kframe}
\end{knitrout}

% 
\fct{eval\_forecasts} and \fct{summarise\_scores} aggregate scores by taking the mean over according to the grouping specified in \code{summarise_by}. In the above example, if \code{summarise_by = c("model", "target_type")}, then scores would be averaged over all other categories to obtain one score per model and forecast target type. For a more detailed analysis we could for example specify \code{summarise_by = c("mode", "target_type", "location")} to additionally stratify by country. Summarised scores can be visualised using the function \fct{scores\_table} as is shown in Figure \ref{fig:score-table}.

If we wanted to have one score per quantile or one per prediction interval range, we could use something like \code{summarise_by = c("model", "quantile", "range")}. This would allow us to examine interval or quantile coverage or makes it possible to analyse the accuracy of the tails of the forecasts. In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.

The user must, however, exercise some caution when aggregating scores. As explained above, many of the metrics are absolute and scale with the magnitude of the quantity to forecast. This makes it sometimes ill-advised to average over them. In the given example, looking at one score per model (i.e. specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by case forecasts, while errors on death forecasts would have little influence. Similarly, aggregating over different forecast horizons is often not a good idea as the mean will be dominated by further ahead forecast horizons.

\begin{figure}[h]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{score_table}\hlstd{(summarised_scores,} \hlkwc{y} \hlstd{=} \hlstr{"model"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{plots/plot-score-table-1} 
\end{knitrout}
\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
\end{figure}

@@

Again there are two possible ways to do pairwise comparisons. First, the function \fct{pairiwse\_comparison} can be called to compare models against each other. Here \code{summarise_by} denotes what relative skill scores shall be computed for. \code{summarise_by = "model"} means there will be one relative skill score per model, while \code{summarise_by = c("target_type", "target_type")} would mean there is one relative skill score per model, computed completely separately for different target types.

When using \fct{pairwise\_comparison}, it is important that the scores used as input have not been summarised in any way.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{pairwise} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(scores,} \hlkwc{baseline} \hlstd{=} \hlstr{"EuroCOVIDhub-baseline"}\hlstd{)}
\hlstd{> }
\hlstd{> }\hlstd{pairwise2} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(scores,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
\hlstd{+ }    \hlstr{"target_type"}\hlstd{),} \hlkwc{baseline} \hlstd{=} \hlstr{"EuroCOVIDhub-baseline"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

@@

\begin{figure}[h!]
\centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{plot_pairwise_comparison}\hlstd{(pairwise2)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{target_type,}
\hlstd{+ }    \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{plots/plot-pairwise-plot-1} 
\end{knitrout}
\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets. If a tile is blue, then the model on the y-axis performed better. If it is red, the model on the x-axis performed better in direct comparison. }
\end{figure}
